A Test of the Frequency Independence Assumption of Power System Coefficients used in Geomagnetically Induced Current Estimates

### 2019-07-24 paper.tex
Updates from Overleaf

\section{Results}

@@

The average coefficients and 95\% CI for Model 1 are $a_o = 73$ $[62,83]$ A/(V/km) and $b_o = -72.0$ $[-98,-45]$ A/(V/km) and the distribution of the coefficients for the segments is shown in Figure~\ref{histogram}. (Results are similar if 1-minute averages are used; \cite{Watari2015} found $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14--2006-12-15 JST time interval).

@@

Figure~\ref{Z} shows that the frequency domain transfer functions for Model~2 vary by an order of magnitude on timescales of 60~s to 12~h (that for Model~1 is constant by definition). For this system, GIC is more sensitive to $E_y$ than $E_x$ due to $b(\omega) > a(\omega)$, but note that $E_y(t)$ has less variability than $E_x(t)$ - the variance in $E_x(t)$ is $\sim$5x larger than $E_y(t)$ over the entire dataset (this is also visible in Figure~\ref{sample}).

@@

(3) The equation of Model~1 assumes that the electric field is spatially uniform at all points in the systems. (An equation of this form can be computed in principle for an electric field with an arbitrary spatial distribution but most often the electric field is assumed to be spatially uniform). The exact calculation of GICs in a power system under the quasi-DC assumption requires knowledge of the electric field at all points of the system, and simulations of GIC flow typically start with an assumed electric field along the lines to calculate input voltages \citep{Albertson1981,Lehtinen1985, Horton2012}. The voltages are then used along with the DC system component resistances and their connectivity to calculate GIC in the different parts of the system. In both Model~1 and Model~2, the input electric field is assumed to be spatially uniform along the line in which GIC was measured so that the potential difference is directly proportional to the electric field. Spatial variations in the electric field measured on Earth's surface can be due to horizontal variations in the incident electric field and horizontal variations in sub-surface conductivity, including those at land-ocean interfaces. If a spatial variation exists along the line in which GIC was measured, the coefficients derived from the data in both models may differ significantly from those calculated using system configuration and resistance information. However, the data-derived coefficients will not have a frequency dependence that results from a variation in the direction of the electric field along the line.

@@

@@

Although the use of frequency independent $a$ and $b$ parameter values estimated from data give a reasonable out-of-sample prediction model, it is not known how much data-estimated parameter values will differ from those directly calculated. (Because the details of the power system under consideration in this work are not available, we were not able to make this comparison.) Works that use Model~1 and generally note a high degree of visual consistency between the measured values and predictions using this model (e.g., the top panel of Figure~\ref{predictions}). However, we have shown using contemporaneous GIC and electric field measurements that a significantly better estimation of GIC can be obtained by using a frequency dependent model of the form of Model~2.

>>>>

\section{Results}
\label{results}

@@

The average coefficients and 95\% CI for Model 1 are $\overlin{a}_o = 73.0$ $[62,83]$ A/(V/km) and $\overline{b}_o = -72.0$ $[-98,-45]$ A/(V/km) and the distribution of the coefficients for the segments is shown in Figure~\ref{histogram}. (Results are similar if 1-minute averages are used; \cite{Watari2015} found $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14 to 2006-12-15 JST time interval).

Note that the values of $\overline{a}_o = 73$ A/(V/km) and $\overline{b}_o = -72$ A/(V/km) imply that for a given electric field magnitude, GIC is largest at the measurement site when the electric field points in the northwest direction (\cite{Pulkkinen2010} and \cite{Watari2015} also found values of $a_o>0$ and $b_o<0$). However, as shown in Figure~\ref{map}, the 187 kV line which terminates at the Memanbetsu substation runs in the northeast direction, and hence the dominant contribution to the measured GIC would be from currents flowing in this line. If the electric field is the same at all points on the line, GIC in the grounded transformers at Memanbetsu substation should be the largest when the electric field points in the northeast direction.

@@

Figure~\ref{Z} shows that the frequency domain transfer functions for Model~2 vary by an order of magnitude on timescales of 60~s to 12~h (that for Model~1 is constant by definition). For this system, GIC is more sensitive to $E_y$ than $E_x$ due to $b(\omega) > a(\omega)$, but note that $E_y(t)$ has less variability than $E_x(t)$ - the variance in $E_x(t)$ is $\sim$5x larger than $E_y(t)$ over the entire dataset (this is also visible in Figure~\ref{sample}). Note that transfer function shown in Figure~\ref{Z} differs significantly from that of a layered slab conductivity model for which $|Z_{xy}|=|Z_{yx}|$.

@@

(3) The equation of Model~1 assumes that the electric field is spatially uniform at all points in the systems. (An equation of this form can be computed in principle for an electric field with an arbitrary spatial distribution but most often the electric field is assumed to be spatially uniform). The exact calculation of GICs in a power system under the quasi-DC assumption requires knowledge of the electric field at all points of the system, and simulations of GIC flow typically start with an assumed electric field along the lines to calculate input voltages \citep{Albertson1981,Lehtinen1985, Horton2012}. The voltages are then used along with the DC system component resistances and their connectivity to calculate GIC in the different parts of the system. In both Model~1 and Model~2, the input electric field is assumed to be spatially uniform along the line in which GIC was measured so that the potential difference is directly proportional to the electric field. Spatial variations in the electric field measured on Earth's surface can be due to horizontal variations in the incident electric field and horizontal variations in sub-surface conductivity, including those at land-ocean interfaces. If a spatial variation exists along the line in which GIC was measured, the coefficients derived from the data in both models may differ significantly from those calculated using system configuration and resistance information. However, the data-derived coefficients will not have a frequency dependence that results only from the variation in the direction of the electric field along the line.

@@

Note that in this discussion, we have attempted to address independently the contribution of spatial variations in the electric field magnitude and direction and spatial variations in the electric field spectra, but generally these variations do not occur independently as they are both caused by horizontal variations in the ground conductivity. Thus, the finding that the angle of the electric field that produces maximum GIC differs significantly from the angle of the 187~kV power line (for both Model~1 and Model~2) is likely to be consistent with the frequency dependence of $a$ and $b$ found in Model~2. Testing this claim would require a full network analysis an knowledge of the ground conductivity throughout the network.

%We have argued above that a spatial variation in the spectra of the electric field can explain the frequency dependence of $a$ and $b$. There is also evidence that the electric field varies along

%A consideration of the ratio of $b_o/a_o$ suggests that the ground conductivity varies significantly along the length of the line. As shown in Figure~\ref{map}, the 187~kV line runs in a direction that is approximately 45$^o$ east of north.  Assuming negligible GIC contributions from other unknown parts of the network and an electric field that is uniform along the length of the line, the equation for Model~1, $GIC(t) = a_oE_x(t) + b_oE_y(t)$, where $x$ is northward and $y$ is eastward,

%one would expect the values of $a_o$ and $b_o$ to be such that $tan^{1}b_o/a_o$ is approximately 45$^o$ east of north .

@@

Although the use of frequency independent $a$ and $b$ parameter values estimated from data give a reasonable out-of-sample prediction model, it is not known how much data-estimated parameter values will differ from those directly calculated. (Due to the limited availability of details of the power system and the ground conductivity along the line, we were not able to make this comparison.) Works that use Model~1 and generally note a high degree of visual consistency between the measured values and predictions using this model (e.g., the top panel of Figure~\ref{predictions}). However, we have shown using contemporaneous GIC and electric field measurements that a significantly better estimation of GIC can be obtained by using a frequency dependent model of the form of Model~2.
